\cleardoublepage

\chapter*{Abstract\markboth{Abstract}{Abstract}}
\noindent Robotics is the discipline focused on the design, construction and programming of robots, machines capable of performing tasks autonomously or controlled by humans. 
autonomously or controlled by humans. Within this, we can find the branch of educational robotics, which is characterised by 
offering robots that are useful for learning. Another branch is 
low-cost robotics, a field of study that seeks to overcome the economic barriers of traditional robots in order to 
economic barriers presented by traditional robots to ensure that everyone can have access to this technology. 
can have access to this technology. 

Robots commonly used in universities are really expensive, so it is difficult to acquire a large number of them. 
it is difficult to acquire a large number of them, limiting their availability. This is why 
This is why a solution to this problem has been developed in this work. Specifically, we have developed 
a low-cost industrial robotic arm that can be manufactured using any conventional 3D printer.

The 3D model has been developed with computer design tools, such as FreeCAD, a powerful 3D modelling application of code 
a powerful open source 3D modelling application that has made it possible to create each of the parts that make up the robotic arm.

On the other hand, the hardware used is easily available on the internet, which has allowed the robot to be developed in an accessible and economical way. 
the development of the robot in an accessible and economical way. The choice of cheap, quality components available on the market 
on the market has been fundamental in guaranteeing its success.

The final robot has been integrated into ROS 2 Humble, an open source middleware widely used in robotic applications. 
widely used in robotic applications. This integration has provided a standardised and efficient interface 
standardised and efficient interface so that any application developed in this ecosystem can make use of it. In addition, it has been configured to be used in the 
has been configured for use in the MoveIt 2 framework, which makes it even easier to use.

With the combination of these tools and technologies, it has been possible to develop a highly functional and robust robot, ready to be used in the field. 
functional and robust robot, ready to be used in an academic environment training new generations. The 
The result represents a significant research and development effort, which opens the door to future improvements and advances in the 
improvements and advances in the field of robotics.

Finally, numerous tests have been carried out to evaluate the different technical aspects that define the performance and characteristics of the robot. 
define the performance and characteristics of the final robot.
